---
title: "Tutorial 4 - Programming Basics 2"
tutorial:
  id: "com.example.tutorials.basics2"
  version: 1.1.0
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to break things and put them back together
---

```{r setup, include=FALSE}
library(learnr)
tutorial_options(exercise.timelimit = 60) # prevent overly complex code from running
knitr::opts_chunk$set(echo = FALSE)
```

# Programming Basics, Part II

## Working with indices 

*As with before, work through these exercises to learn more about the concepts just covered. There has been a lot we've covered so hopefully this will solidify it a little bit better.*  

*Reminder: clicking "Run Code" evaluates everything in the box. This is slightly different than in RStudio where we usually run things line by line.*


### Exercise 1 {data-allow-skip=TRUE}

Use square brackets `[ ]` to select the third element of `x` below.

```{r x1, exercise=TRUE, exercise.lines = 5}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)
x
```

```{r x1-solution}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)
x[3]
```

### Exercise 2 {data-allow-skip=TRUE}

*What happens when there are multiple indices used?*

Use square brackets `[ ]` to select the third, fourth, and fifth element of `x` below.

```{r x2, exercise=TRUE, exercise.lines = 5}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)
x
```

<div id="x2-hint">
**Hint:** Construct a vector _inside_ the square brackets with `c(...)` or `:`
</div>

```{r x2-solution}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)
x[c(3, 4, 5)]
```

### Exercise 3 {data-allow-skip=TRUE}

*Storing numbers to a subset?  You can do that??*

Use the `<-` operator to store `999` to the third element of `x` below.  Add another line below, printing `x` to see what happened to its contents.

```{r x3, exercise=TRUE, exercise.lines = 5}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)

x[3]
```

```{r x3-solution}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)

x[3] <- 999

x
```

### Exercise 4 {data-allow-skip=TRUE}

*What happens when there are multiple indices used?*

Use the `<-` operator to store `999` to the third, fourth, and fifth element of `x` below.  Add another line below, printing `x` to see what happened to its contents.

```{r x4, exercise=TRUE, exercise.lines = 5}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)

x[c(3, 4, 5)] 
```

```{r x4-solution}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)

x[c(3, 4, 5)] <- 999

x
```


## Sticking things together

### Exercise 5 {data-allow-skip=TRUE}

*Row-wise*

Use `rbind()` to stick vectors `a` and `b` together as multiple rows.

```{r x5, exercise=TRUE, exercise.lines = 5}
a <- 1:5
b <- 6:10
```

```{r x5-solution}
a <- 1:5
b <- 6:10
rbind(a, b)
```

### Exercise 6 {data-allow-skip=TRUE}

*Column-wise*

Now, use `cbind()` to stick vectors `a` and `b` together as multiple columns!

```{r x6, exercise=TRUE, exercise.lines = 5}
a <- 1:5
b <- 6:10
```

```{r x6-solution}
a <- 1:5
b <- 6:10
cbind(a, b)
```

### Exercise 7 {data-allow-skip=TRUE}

*Length-wise*

What happens when use `a` and `b` as inputs in the `c()` function?

```{r x7, exercise=TRUE, exercise.lines = 5}
a <- 1:5
b <- 6:10
```

```{r x7-solution}
a <- 1:5
b <- 6:10
c(a, b)
```


## Working with two dimensions!

### Exercise 8 {data-allow-skip=TRUE}

```{r loaddata, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
exampledf <- data.frame(sex = c("Male", "Male", "Female", "Male", "Female", "Female", "Male", "Male", "Female", "Male", "Female", "Male", "Female"), 
                        length = c(110, 112, 90, 89, 107, 104, 98, 102, 92, 88, 102, 109, 99), 
                        weight = c(3, 3.4, 2.4, 2.5, 3.0, 2.9, 2.8, 2.8, 2.3, 2.3, 2.7, 3.1, 2.7),
                        age = c(2, 2, 1, 1, 2, 2, 2, 2, 1, 1, 2, 2, 2),
                        statarea = c(115, 115, 101, 115, 115, 101, 115, 115, 101, 101, 115, 101, 101))

```

*Here comes a data.frame...*

A data.frame has been loaded to the environment, and the first few rows are printed below.  The `head()` function can be handy for getting a quick look at a data.frame, without printing the whole thing!  Change the argument `n=` to show the first ten rows instead.

```{r x8, exercise=TRUE, exercise.lines = 5, exercise.setup="loaddata"}
head(exampledf, n=5)
```

```{r x8-solution}
head(exampledf, n=10)
```

### Exercise 9 {data-allow-skip=TRUE}

*Another way of looking at the structure of our data.frame*

The `str()` function is a bit more flexible with regard to input, and prints the structure of an object.  Change `head` to `str` in the code below and see what it does.

```{r x9, exercise=TRUE, exercise.lines = 5, exercise.setup="loaddata"}
head(exampledf)
```

```{r x9-solution}
str(exampledf)
```

### Exercise 10 {data-allow-skip=TRUE}

*Extracting columns - one way*

Notice that when we called `str()`, the columns names were preceded by the `$` character.  That means we can use `$` to grab a column at a time!  Add a line to the code below to extract the `length` column.

```{r x10, exercise=TRUE, exercise.lines = 5, exercise.setup="loaddata"}
str(exampledf)
```

<div id="x10-hint">
**Hint:** `exampledf$`
</div>

```{r x10-solution}
exampledf$length
```

### Exercise 11 {data-allow-skip=TRUE}

*Extracting columns - another way!*

We can also use row or column indices inside square brackets `[]`.  Change the second line below to extract length, which is the second column of the data.frame.

```{r x11, exercise=TRUE, exercise.lines = 5, exercise.setup="loaddata"}
exampledf$length

exampledf[]
```

<div id="x11-hint">
**Hint:** `exampledf[,]`
</div>

```{r x11-solution}
exampledf$length

exampledf[, 2]
```

### Exercise 12 {data-allow-skip=TRUE}

*Extracting rows*

Modify the code below to extract the third row of the data.frame.

```{r x12, exercise=TRUE, exercise.lines = 5, exercise.setup="loaddata"}
exampledf[]
```

```{r x12-solution}
exampledf[3, ]
```

### Exercise 13 {data-allow-skip=TRUE}

*Extracting multiple rows??*

Modify the code below to use a vector input to extract _rows three through five_ of the data.frame.  How do you think you would do this?

```{r x13, exercise=TRUE, exercise.lines = 5, exercise.setup="loaddata"}
exampledf[3, ]
```

<div id="x13-hint">
**Hint:** `c()` or maybe `:` ...
</div>

```{r x13-solution}
exampledf[3:5, ]
```

### Exercise 14 {data-allow-skip=TRUE}
*Summarizing a data.frame*

The `summary()` function can be a useful tool for summarizing many kinds of objects.  Run the code below to show a summary of the `$length` column.  Then, add a line below it to summarize the whole data.frame at once.

```{r x14, exercise=TRUE, exercise.lines = 5, exercise.setup="loaddata"}
summary(exampledf$length)
```

<div id="x14-hint">
**Hint:** You can pass almost anything to the `summary()` function - including data.frames!
</div>

```{r x14-solution}
summary(exampledf$length)

summary(exampledf)
```

### Exercise 15 {data-allow-skip=TRUE}
*When summarizing a data.frame shows off its problems!*

Uh oh - it looks like stat area is formatted as a number!  That's probably not what we want!  Good thing we can fix it.  Turn it into a factor with `as.factor(exampledf$statarea)`, and store it to `exampledf$statarea`.  Add a line below it to re-run `summary()` and see what changed.

Note: for completeness, we should probably do the same thing to sex, which is currently formatted as character.

```{r x15, exercise=TRUE, exercise.lines = 5, exercise.setup="loaddata"}
summary(exampledf)
```

<div id="x15-hint">
**Hint:** exampledf$statarea <- 
</div>

```{r x15-solution}
summary(exampledf)

exampledf$statarea <- as.factor(exampledf$statarea)

summary(exampledf)
```

### Exercise 16 {data-allow-skip=TRUE}
*Extracting pieces of a list*

Your colleague ran a two-sample t-test to compare the mean lengths from stat areas 101 and 115, and saved the results in an object called `t_test_results`.  Run the line below to see how R will print the results of a two sample t-test.

You'd actually like to pull the 95% confidence interval of the difference in means so you can use it own code, without a lot of copy-pasting!  First, add a line that calls `str()` to look at the structure of the test results, and re-run your code.

Then, add another line to extract the CI.

Bonus: How would you extract just the _lower bound_ of the CI?

```{r testresults, exercise.setup="loaddata"}
testresults <- with(exampledf, t.test(length[statarea=="101"], length[statarea=="115"]))
```

```{r x16, exercise=TRUE, exercise.lines = 5, exercise.setup="testresults"}
testresults
```

<div id="x16-hint">
**Hint:** `testresults$` 
</div>

```{r x16-solution}
str(testresults)

testresults$conf.int

testresults$conf.int[1]
```

## Fun with logicals

### Exercise 17 {data-allow-skip=TRUE}

*Subsetting a data.frame*

Modify the code below to print _only_ observations where `sex` is equal to `"Female"`.

```{r loaddata2, exercise.setup="loaddata"}
exampledf$statarea <- as.factor(exampledf$statarea)
exampledf$sex <- as.factor(exampledf$sex)
```

```{r x17, exercise=TRUE, exercise.lines = 5, exercise.setup="loaddata2"}
exampledf[exampledf$sex=="", ]
```

```{r x17-solution}
exampledf[exampledf$sex=="Female", ]
```

### Exercise 18 {data-allow-skip=TRUE}

*Subsetting a data.frame*

Modify the code below to print _only_ observations where `sex` is equal to `"Female"` _and_ `age` is equal to `2`.

Note that the `&` character means "and", and the `|` character means "or".

```{r x18, exercise=TRUE, exercise.lines = 5, exercise.setup="loaddata2"}
exampledf[, ]
```

<div id="x18-hint">
**Hint:** Did you use `exampledf$` ?
</div>

```{r x18-solution}
exampledf[exampledf$sex=="Female" & exampledf$age==2, ]
```

### Exercise 19 {data-allow-skip=TRUE}

*Same same, but different*

Modify the code below to print observations where `sex` is equal to `"Female"` _or_ `statarea` is equal to `"115"`.

Note that the `&` character means "and", and the `|` character means "or".

```{r x19, exercise=TRUE, exercise.lines = 5, exercise.setup="loaddata2"}
exampledf[, ]
```

```{r x19-solution}
exampledf[exampledf$sex=="Female" | exampledf$statarea=="115", ]
```

### Exercise 20 {data-allow-skip=TRUE}

*Making a new column and using ifelse()*

One way to create a new column in our example data.frame is by assigning a new vector of data to a new column name of your choice.

Create a new column called "size", and populate it with `"Small"` if `exampledf$length < 100`, and with `"Big"` otherwise, using the `ifelse()` function.

Feel free to consult `?ifelse` if you'd like!

```{r x20, exercise=TRUE, exercise.lines = 5, exercise.setup="loaddata2"}
exampledf$size <- 
  
exampledf
```

```{r x20-solution}
exampledf$size <- ifelse(exampledf$length < 100, "Small", "Big")
  
exampledf
```

### Exercise 21 {data-allow-skip=TRUE}

*Final challenge: dealing with NA values*

```{r loaddata3, exercise.setup="loaddata2"}
exampledf$sex[c(1, 4, 9)] <- NA
```

Oops!  It turns out there was something wrong with some of the samples, and some of the entries for sex needed to be redacted.  Create a new column that says `"Good sample"` if `!is.na(exampledf$sex)` and `"Bad sample"` otherwise.

```{r x21, exercise=TRUE, exercise.lines = 5, exercise.setup="loaddata3"}
exampledf$goodsample <- 
  
exampledf
```

<div id="x21-hint">
**Hint:** Did you use `ifelse()` ?
</div>

```{r x21-solution}
exampledf$goodsample <- ifelse(!is.na(exampledf$sex), "Good sample", "Bad sample")
  
exampledf
```


### COMPLETE
Whew!  That section was a monster.  Great work!