---
title: "Tutorial 4 - Programming Basics 2"
tutorial:
  id: "com.example.tutorials.basics2"
  version: 1.0.0
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to break things and put them back together
---

```{r setup, include=FALSE}
library(learnr)
tutorial_options(exercise.timelimit = 60) # prevent overly complex code from running
knitr::opts_chunk$set(echo = FALSE)
```

# Programming Basics, Part II

## Working with indices 

### Exercise 1 {data-allow-skip=TRUE}

*Use this space to work through code to understand the concepts we've gone over. Go slow and try to get the correct answers. This is not graded and is not a race, use this space to familiarize yourself with how R works. Once you are finished, just close out of this before opening a new tutorial.*
*This mini-console is running R, however clicking "Run Code" evaluates everything written in there; in RStudio, we run through it line-by-line.*

Use square brackets `[ ]` to select the third element of `x` below.

```{r x1, exercise=TRUE, exercise.lines = 5}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)
x
```

```{r x1-solution}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)
x[3]
```

### Exercise 2 {data-allow-skip=TRUE}

*What happens when there are multiple indices used?*

Use square brackets `[ ]` to select the third, fourth, and fifth element of `x` below.

```{r x2, exercise=TRUE, exercise.lines = 5}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)
x
```

<div id="x2-hint">
**Hint:** Construct a vector _inside_ the square brackets with `c(...)` or `:`
</div>

```{r x2-solution}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)
x[c(3, 4, 5)]
```

### Exercise 3 {data-allow-skip=TRUE}

*Storing numbers to a subset?  You can do that??*

Use the `<-` operator to store `999` to the third element of `x` below.

```{r x3, exercise=TRUE, exercise.lines = 5}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)

x[3]

x
```

```{r x3-solution}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)

x[3] <- 999

x
```

### Exercise 4 {data-allow-skip=TRUE}

*What happens when there are multiple indices used?*

Use the `<-` operator to store `999` to the third, fourth, and fifth element of `x` below.

```{r x4, exercise=TRUE, exercise.lines = 5}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)

x[c(3, 4, 5)] 

x
```

```{r x4-solution}
x <- c(1, 1, 2, 3, 5, 8, 13, 21)

x[c(3, 4, 5)] <- 999

x
```


## Sticking things together

### Exercise 5 {data-allow-skip=TRUE}

*Row-wise*

Use `rbind()` to stick vectors `a` and `b` together as multiple rows.

```{r x5, exercise=TRUE, exercise.lines = 5}
a <- 1:5
b <- 6:10
```

```{r x5-solution}
a <- 1:5
b <- 6:10
rbind(a, b)
```

### Exercise 6 {data-allow-skip=TRUE}

*Column-wise*

Now, use `cbind()` to stick vectors `a` and `b` together as multiple columns!

```{r x6, exercise=TRUE, exercise.lines = 5}
a <- 1:5
b <- 6:10
```

```{r x6-solution}
a <- 1:5
b <- 6:10
cbind(a, b)
```

### Exercise 7 {data-allow-skip=TRUE}

*Length-wise*

What happens when we put vectors in the `c()` function

```{r x7, exercise=TRUE, exercise.lines = 5}
a <- 1:5
b <- 6:10
```

```{r x7-solution}
a <- 1:5
b <- 6:10
c(a, b)
```


## Working with indices in two dimensions!


str with a data.frame

$ with a data.frame

summary with a vector
summary with a data.frame
turn something in to a factor, then summary again

str with a list

$ with a list


logicals

logicals to subset stuff

is.na, !is.na, ifelse