---
title: "Tutorial 3 - Programming Basics 1"
tutorial:
  id: "com.example.tutorials.basics1"
  version: 1.1.1
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to math in R.
---

```{r setup, include=FALSE}
library(learnr)
tutorial_options(exercise.timelimit = 60) # prevent overly complex code from running
knitr::opts_chunk$set(echo = FALSE)
```

# Programming Basics, Part I  
## Vectors and vector math 

### Exercise 1 {data-allow-skip=TRUE}

*Use this space to work through code to understand the concepts we've gone over. Go slow and try to get the correct answers. This is not graded and is not a race, use this space to familiarize yourself with how R works.*  
*If you don't finish this, that's totally fine! Come back and finish this later.*  
*This mini-console is running R, however clicking "Run Code" evaluates everything written in there; in RStudio, we run through it line-by-line.*

Modify the code below to create a vector of numbers from 1 to 5, then hit "Run Code" (or ctrl+enter):

```{r myvector, exercise=TRUE, exercise.lines = 5}
myvector <- c(, , , , )
myvector
```

```{r myvector-solution}
myvector <- c(1, 2, 3, 4, 5)
myvector
```

### Exercise 2

*Vector math, part I*

Modify the second line of the code below to multiply all elements of `myvector` by 10.

```{r vector1, exercise=TRUE, exercise.lines = 5}
myvector <- c(1, 2, 3, 4, 5)
myvector
```


```{r vector1-solution}
myvector <- c(1, 2, 3, 4, 5)
myvector * 10
```

### Exercise 3

*Vector math, part II: this time it's personal*

Create a new variable called `CPUE` (catch per unit effort), defined by `catch` divided by `hours`.  Print the result to the console!

```{r vector2, exercise=TRUE, exercise.lines = 5}
catch <- c(33, 78, 12, 45, 89)
hours <- c(1.5, 6.7, 5, 8.2, 10)
```

<div id="vector2-hint">
**Hint:** `CPUE <- `
</div>


```{r vector2-solution}
catch <- c(33, 80, 12, 45, 89)
hours <- c(1, 4, 5, 8.2, 10)

CPUE <- catch / hours
CPUE
```

## More vectors!

### Exercise 4

*Regular sequences*

You can use the `:` character to create a regular sequence, such as `1:10`.  Most mathematical operators (such as `sqrt()`, or square root) will accept a vector input.

Modify the code below to compute the square root of all whole numbers from 1 to 50.  Does element 49 have a value of 7.000?

```{r vector3, exercise=TRUE, exercise.lines = 5}
sqrt()
```

```{r vector3-solution}
sqrt(1:50)
```

### Exercise 5

*Regular sequences, customized!*

The `seq()` function can be used to generate sequences using optional `from=`, `to=`, `by=`, and `length.out=` arguments.

Replace the `99` in each line below so that all lines have the same output!  

```{r vector4, exercise=TRUE, exercise.lines = 10}
c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100)

seq(from=0, to=100, by=99)

seq(from=0, to=100, length.out=99)

(0:10)*99  # bonus!
```

```{r vector4-solution}
c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100)

seq(from=0, to=100, by=10)

seq(from=0, to=100, length.out=11)

(0:10)*10  # bonus!
```

## Functions that expect vectors

### Exercise 6

*Summary statistics*

Modify the code below to compute the mean, median, and standard deviation of `y`.  

```{r vector5, exercise=TRUE, exercise.lines = 10}
y <- c(1, 1, 2, 3, 5, 8, 13, 21)

mean()

median()

sd()
```

```{r vector5-solution}
y <- c(1, 1, 2, 3, 5, 8, 13, 21)

mean(y)

median(y)

sd(y)
```

### Exercise 7

*Bug hunter!*

This one is tricky!  Run the code below.  Does it give the answer you would expect?  

Modify it to give the _actual_ arithmetic mean of the vector of numbers (6.75).

Hint: `mean()` is expecting a vector input...

```{r vector7, exercise=TRUE, exercise.lines = 10}
mean(1, 1, 2, 3, 5, 8, 13, 21)
```

<div id="vector7-hint">
**Hint:** `c(...)`
</div>

```{r vector7-solution}
mean(c(1, 1, 2, 3, 5, 8, 13, 21))
```

### Exercise 8

*Dealing with NA*

Evaluate the code below.  What happens?

Investigate the `median()` function using `?median`.  How would you compute the sample median without including the `NA` value?

```{r vector6, exercise=TRUE, exercise.lines = 10}
x <- c(1, 1, 2, 3, 5, 8, NA, 21)
median(x)
```

<div id="vector6-hint">
**Hint:** `na.rm=`
</div>

```{r vector6-solution}
x <- c(1, 1, 2, 3, 5, 8, NA, 21)
median(x, na.rm=TRUE)
```




## Quiz 

*Let's review some of the concepts we've gone over.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz-basics1}
quiz(
  question("Which function would you use for taking the mean / average?",
    answer("average()"),
    answer("tools"),
    answer("mean()", correct = TRUE),
    answer("codetools")
  ),
  question("How would you set variable 'fishcolor' to be equal to red, blue, and green?",
    answer("fishcolor <- red, blue, green "),
    answer("fishcolor <- c('red', 'blue', 'green')", correct = TRUE),
    answer("fishcolor <- c(red, blue, green)"),
    answer("'fishcolor' <- c('red, blue, green')")
  ),
  question("How would you get help and learn more about a function named cbind()?",
    answer("help(cbind)", correct = TRUE),
    answer("Search cbind in help panel on bottom-right of RStudio", correct = TRUE),
    answer("Google search 'r cbind()'", correct = TRUE),
    answer("Call 867-5309 and ask for cbind"),
    answer("?cbind", correct = TRUE)
  ),
  question("Select the only line that does NOT have an issue",
    answer("mean(c(3, 4, 5, 6)"),
    answer("mean(2, 3, 4, 5)"),
    answer("mean(c(3, 4, 5, 6))", correct = TRUE),
    answer("mean(x"),
    answer("mean((x)")
  ),
  question("What is the expected outcome of this code: mean(c(2, 2, 4, 6, 6, NA, 2, 6, 5, 3, 4))",
    answer("NA", correct = TRUE),
    answer("4")
  ),
  question("Which of the following creates a vector of numbers?",
    answer("makelist(1, 2, 3)"),
    answer("str(1, 2, 3)"),
    answer("listed(1, 2, 3)"),
    answer("c(1, 2, 3)", correct = TRUE)
  ),
  question("While taking the mean, your answer is returned as NA. What argument should you add inside the function mean()?",
    answer("completecase = FALSE"),
    answer("na.rm = TRUE", correct = TRUE),
    answer("pleaseremoveNAs")
  ),
  question("Oh no! Your console shows a '+' and appears to be waiting for you. What caused this and how to you get it back to normal?",
    answer("Cause: Code is running. Solution: Wait. Wait some more. Wait again?"),
    answer("Cause: Incorrect voodoo chant. Solution: Burn computer"),
    answer("Cause: Wrong function name. Solution: Never make a mistake ever again"),
    answer("Cause: Incomplete code ending. Solution: Press Escape", correct = TRUE)
  ),
  question("Which of these is a good variable name",
    answer("x"),
    answer("sablefishcatch2019", correct = TRUE),
    answer("mean"),
    answer("vector"),
    answer("dat"),
    answer("1sablefish")
  )

)
```

###   

### Quiz Review
How did you do?!  If you've gotten this far, you're doing quite well! Here are some notes about _why_ those were the correct answers:   
 - Knowing when to use quotations feels quite odd at first. There is indeed a method to the madness: items within quotes are usually "text".  
 - If you see the plus sign (+) in the console, R is waiting for you and you probably forgot a closing parenthesis.  
 - Proper variable naming is often a personal choice but the bad ones shown here were all poor choices. Don't use variable names that are extremely vague, have the same wording as a common function, or start with a number.   




### COMPLETE
Great work! You've successfully completed this lesson!
