---
title: "Tutorial 2 - Introducing R and RStudio"
tutorial:
  id: "com.example.tutorials.introducingr"
  version: 1.0.0
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn the basics of how R works.
---

```{r setup, include=FALSE}
library(learnr)
tutorial_options(exercise.timelimit = 60) # prevent overly complex code from running
knitr::opts_chunk$set(echo = FALSE)
```

# Introducing R and RStudio  
## Basic R Programming 

### Exercise 1 {data-allow-skip=TRUE}

*Use this space to work through code to understand the concepts we've gone over. Go slow and try to get the correct answers. This is not graded and is not a race, use this space to familiarize yourself with how R works. Once you are finished, just close out of this before opening a new tutorial.*
*This mini-console is running R, however clicking "Run Code" evaluates everything written in there; in RStudio, we run through it line-by-line.*

In the space below, write `2 + 2`, then hit "Run Code" (or ctrl+enter):

```{r twoplus, exercise=TRUE, exercise.lines = 5}

```

```{r twoplus-solution}
2 + 2
```

### Exercise 2

*We can use the `<-` operator to store a variable.*

Add a second line to the code below, adding one to variable `x`.

```{r store-x, exercise=TRUE, exercise.lines = 5}
x <- 3
```


```{r store-x-solution}
x <- 3
x + 1
```

### Exercise 2 - Review
Reassuring, no?



### Exercise 3 {data-allow-skip=TRUE}

*R can also store text using quotes.*

In the space below, store your name to variable `MyName` by writing it between the quotes.  What do you think you
need to do to tell R to print the contents of `MyName`?

```{r store-name, exercise=TRUE, exercise.lines = 5}
MyName <- ""
```

<div id="make-vector-hint">
**Hint:** How did we print the contents of `x + 1` in the last exercise?
</div>

```{r store-name-solution}
MyName <- "Ozymandias"
MyName
```



### Exercise 4 {data-allow-skip=TRUE}

*Adding comments and "commenting out" lines*

R will not evaluate code to the right of a `#` character.

Run the code below and see what happens.  Make a change to the text comment and see what happens (hopefully nothing!)

How do you think you can tell R to ignore this line altogether?

```{r comments, exercise=TRUE, exercise.lines = 5}
1 + 2 - 3 * 4 / 5   # here is an amazing comment
```

```{r comments-solution}
# 1 + 2 - 3 * 4 / 5   # here is an amazing comment
```