---
title: "Tutorial 2 - Introducing R and RStudio"
tutorial:
  id: "com.example.tutorials.introducingr"
  version: 1.1.0
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn the basics of how R works.
---

```{r setup, include=FALSE}
library(learnr)
tutorial_options(exercise.timelimit = 60) # prevent overly complex code from running
knitr::opts_chunk$set(echo = FALSE)
```

# Introducing R and RStudio  
## Basic R Programming 

### Exercise 1 {data-allow-skip=TRUE}

*Use this space to work through code to understand the concepts we've gone over. Go slow and try to get the correct answers. This is not graded and is not a race, use this space to familiarize yourself with how R works. Once you are finished, just close out of this before opening a new tutorial.*
*This mini-console is running R, however clicking "Run Code" evaluates everything written in there; in RStudio, we run through it line-by-line.*

In the space below, write `2 + 2`, then hit "Run Code" (or ctrl+enter):

```{r twoplus, exercise=TRUE, exercise.lines = 5}

```

```{r twoplus-solution}
2 + 2
```

### Exercise 2

*We can use the `<-` operator to store a variable.*

Add a second line to the code below, adding one to variable `x`.

```{r store-x, exercise=TRUE, exercise.lines = 5}
x <- 3
```


```{r store-x-solution}
x <- 3
x + 1
```

### Exercise 2 - Review
Reassuring, no?



### Exercise 3 {data-allow-skip=TRUE}

*R can also store text using quotes.*

In the space below, store your name to variable `MyName` by writing it between the quotes.  What do you think you
need to do to tell R to print the contents of `MyName`?

```{r store-name, exercise=TRUE, exercise.lines = 5}
MyName <- ""
```

<div id="make-vector-hint">
**Hint:** How did we print the contents of `x + 1` in the last exercise?
</div>

```{r store-name-solution}
MyName <- "Ozymandias"
MyName
```



### Exercise 4 {data-allow-skip=TRUE}

*Adding comments and "commenting out" lines*

R will not evaluate code to the right of a `#` character.

Run the code below and see what happens.  Make a change to the text comment and see what happens (hopefully nothing!)

How do you think you can tell R to ignore this line altogether?

```{r comments, exercise=TRUE, exercise.lines = 5}
1 + 2 - 3 * 4 / 5   # here is an amazing comment
```

```{r comments-solution}
# 1 + 2 - 3 * 4 / 5   # here is an amazing comment
```

## Quiz 

*Let's review some of the concepts we've gone over.*

```{r data-manip-quiz1}
quiz(
  question("Why should you do data manipulation in R rather than Excel?",
    answer("You have a record of what changes were implemented", correct = TRUE),
    answer("Can be easily repeated every time you output from a database / data is updated", correct = TRUE),
    answer("You can leave notes & documentation in the same order as the manipulation steps", correct = TRUE),
    answer("Easier & faster for many tasks", correct = TRUE),
    answer("Because Justin & Matt said to")
  ),
  question("What is the difference between R (R Project) and RStudio?",
    answer("You can customize the layout & appearance of R Project, but not RStudio"),
    answer("No difference in language: RStudio runs R in the background, and provides an organized interface", correct = TRUE),
    answer("RStudio provides easy access to some extra tools", correct = TRUE)
  ),
  question("The best ways to learn R programming are (probably)",
    answer("Try stuff and see what works and what doesn't! (and try to figure out why)", correct = TRUE),
    answer("Read THE R BOOK cover-to-cover before trying ANYTHING"),
    answer("Slip Brian Ripley a $20 and hope he gives you the cheat codes"),
    answer("Start using R for your own work (and be patient at first)", correct = TRUE),
    answer("Connect with other R users", correct = TRUE)
  ),
    question("What happens to variables saved in the Environment when we close R?",
    answer("They're stored to the hard drive"),
    answer("They go away.  Re-run your code to get them back.", correct = TRUE),
    answer("They're gone forever")
  ),
    question("TERMINOLOGY: R Code is evaluated in the ___ pane, but we recommend writing your scripts in the ___ pane so you have a record of what you did",
    answer("Console / Editor", correct = TRUE),
    answer("Editor / Console"),
    answer("Help / Plots"),
    answer("Packages / Workspace")
  ),
  question("What happens if a line of your R code causes an error?",
    answer("Computer explodes. Solution: buy new computer."),
    answer("You will get a message that might give insight as to what went wrong", correct = TRUE),
    answer("You might learn something new about R in diagnosing the problem", correct = TRUE),
    answer("You might find some great resources or helpful people in the R community while troubleshooting", correct = TRUE),
    answer("Hadley Wickham will blacklist you from the tidyverse"),
    answer("Nothing that doesn't happen to Justin & Matt several times a day!", correct = TRUE)
  )
)
```

###

### COMPLETE
Great work! You've successfully completed this lesson!